### 2018-04-28 response.tex

\introcomment{~}{%
In particular, I agree with their comment that the limitations of the study, i.e., the choice of 11 metrics and earthquake simulations for only one earthquake, needs to be included in the abstract and the conclusions.}

\response{%
We modified the abstract and the conclusions accordingly. In the case of the abstract we added minor modifications to emphasize the use of 11 metrics, and a whole new sentence explicitly referencing the use of three simulation datasets corresponding to the 2008 Chino Hills earthquake. In the case of the conclusions, although we already had a paragraph describing the limitations of the study, we added a small sentence reemphasizing it.
}

\introcomment{~}{%
In addition, I also agree that all the terms and methodology should be well defined for the BSSA readership, so that the paper can be understood by them without resort to frequent online searches.
}

\response{%
Thank you for this suggestion. We revised the \textbf{Introduction}, \textbf{Study Dataset}, and \textbf{Data Analysis} sections to carefully introduce and explain the different terms and their equivalencies, when necessary. We think the introduced modifications should solve this issue and provide a long-lasting reference for the BSSA readership.
}

@@

\response{%
We thank the reviewer for the positive feedback in this general review. We agree that there were some important points that we did not discuss fully in the manuscript and we have corrected them. In attention to this general review, we modified the abstract, the introduction, and the conclusions to (i) transparently point out the limitations of our study, and (ii) offer an additional perspective on the potential refinement of it or use with additional datasets or in reference to additional validation metrics.
}

\comment{1}{%
The terminologies used in the paper are not clear in many cases. For example, what are ``features'' ``attributes'' ``dimensions'' ``trees, nodes and leafs'' ``class''? It was very difficult to read the section on Data Analysis Method and Introduction since these terminologies were not clearly defined or described in terms of the variables used in this paper. I was confused what ``data'' and ``results'' were referring to: GOF score data? GOF metric data? GOF category? Something other than GOF? I suggest that the authors streamline the paper and the terminologies to be consistent and clear in the final version.
}

\response{%
We revised the \textbf{Introduction}, \textbf{Study Dataset}, and \textbf{Data Analysis} sections to carefully introduce and explain the different terms and their equivalencies.
}

@@

\comment{4}{%
In introduction, the difference between ``verification'' and ``validation'' should be discussed.
}

\response{%
A brief description of the differences between verification and validation was added to the Introduction.
}

\comment{5}{%
In abstract and introduction, what does a ``direct application'' mean? it's not clear. Similarly, in Validation Metrics section: ``direct quantitative comparisons'' and ``indirect statistical analysis''.
}

\response{%
The term ``direct application'' was present in the Introduction, but it has been removed due to improvements in the text based on other suggestion by this and a Reviewer \#2. The terms ``direct quantitative comparisons'' and ``indirect statistical analysis'' were present in the Introduction and the Validation Metrics sections. In our view, the first of these terms makes reference to validation through direct, i.e., signal to signal, quantitative comparisons of metrics; while the second of these terms makes reference to validation by means of statistical analysis showcasing the overall results of comparisons, and thus the use of ``indirect''. We understand, however, that this may not be a view shared by any other reader, so we have modified the text to be more explicit about this distinction.
}

@@

\comment{11}{%
Page 14, line 323: It's not clear how you ``replicate data'', provide more detail.
}

\response{%
We modified this paragraph to provide more details about the oversampling process.
}

\comment{12}{%
I don't think ``confusion matrix'' given in tables is explained in the paper!
}

\response{%
The explanation about confusion matrices in the original manuscript was in the section titled \textbf{Decision Trees}, page 13, between lines 280 and 282. The explanation included an illustrative example for the simplest case of two categories. We assumed it would be clear that this concept is extendable to larger matrices such as those shown in Tables 2 and 3. We added a sentence in the explanation to make it explicit that they can be extended to deal with more attributes.
}

@@

\comment{14}{%
Page 18, line 405: this is implying that fig 10 is similar to fig2! I personally cannot see this in the figures! They look very different. It might be better to regenerate fig 11a but with different color scales that only vary by four choices, then the comparison would be much easier.
}

\response{%
In hindsight, we may have allowed ourselves to be carried away by our own optimism. We have therefore modified our comments in this section to a more self-critical approach about the results. The question is really not whether they look the same or not, but whether they lead to similar---if not equivalent---conclusions. It is our interpretation that they do in the sense that looking at the results obtained with either approach one may walk away with the same conclusion about the simulation as a whole and about how well (or not well) did the simulations turned out with respect to specific areas in the simulation domain. We modified the text to convey this message, now hopefully more clearly than before.
}

\comment{15}{%
Figure 8: I don't know what you mean by y-label ``decision attributes''. Do you mean ``the number of GOF metrics''? why there is no data point for 1 or 11?
}

\response{%
That is right. It should be the number of decision attributes. We modified the figure accordingly. The reason why there is no data for 1 is because the analysis did not yield any tree that would classify the data using a single metric. On the other end, the reason there are no trees with the 11 metrics is because the analysis of the data resulted in the metric C11 not having any participation in the decision process, as shown in Table 6. We also tried other trees for which we removed metrics C5 and C8. In such cases the participations of the metrics changes and so do the topology of the trees. Some of the trees resulting from that analysis did include the metric C11, but with a very small participation, as shown un the new Table 7.
}

@@

\comment{16}{%
The conclusion basically says that response spectrum is the most important validation metric. This is not surprising to any engineer or modeler. The big question is what other parameters describe the differences between a simulation and a recorded motion if their response spectra matched? Or in other words, what are the other important validation parameters aside from response spectrum (many have argued that's duration, but no statistical proof). Can you use your method to answer this question? Can you condition your machine learning process on Sa being the same (for example use a subset of data with similar Sa or artificially make them equal), then decide what's the next important parameter? Would the result be the same as in your conclusions? Or would it be influenced by removal of Sa as a parameter (this is possible since there are correlations between Sa and other metrics).
}

\response{%
These are very good points. Yes, the general consensus about the importance of the response spectra (C8) may not be surprising, but it is still very useful to confirm such importance through a data-informed process. Most important, as pointed out by the reviewer, is to answer the question about what other parameters are relevant. We think our work contributes precisely to that. We mention in the conclusion that the next metrics in line are the energy (C4), the peak acceleration (C5) and the peak velocity (C6). In order to further emphasize this point, as described in the response to comment \#8, we also included a new table (Table 7) showing the participation of the different metrics in the absence of the response spectra (C8) and the peak acceleration (C5). The results confirm that in such cases the next most relevant metrics are still the energy (C4) and the peak ground velocity (C6), as mentioned before. This additional analysis also reafirms the initial conclusion about the very small (and in some cases null) participation of the duration (C11) in most of the decision trees built in the data-processing analysis carried out here.
}

@@

\introcomment{~}{%
Earthquake ground motion simulations are usually validated against observational data or empirically-derived solutions. However, exactly which parameters to compare (e.g., peak ground velocity, duration, spectral content) in order to assess the goodness-of-fit, is not always clear. This paper uses machine learning techniques and eleven possible metrics for comparing (recorded and simulated) ground motions to determine the parameters that are most predictive of goodness-of-fit. This is an interesting paper, and will help future researchers prioritize which metrics to focus on when validating ground motions simulations. The paper is generally well written.
}

\response{%
We thank the reviewer for the generally positive view of our paper.
}

\introcomment{~}{%
As a strong ground motion seismologist (i.e., I assume I'm part of the audience you hope to reach), there is a lot of machine learning-related jargon that makes certain sections difficult to follow.
}

\response{%
Thank you for this suggestion. We revised the \textbf{Introduction}, \textbf{Study Dataset}, and \textbf{Data Analysis} sections to carefully introduce and explain the different terms and their equivalencies, when necessary.
}

@@

\introcomment{~}{%
It would also be useful to have some discussion regarding which parameters were found to be repetitive, in case future authors wish to look at more than the four recommended parameters.
}

\response{%
We addressed this by considering additional trees in the cases of removing metrics C5 and C8 (see Table 7). This was also done in response to the comments \#8 and \#16 from Reviewer 1. As we explain in the response to those comments, the procedure used here is different from a simpler correlation analysis in which a metric can be directly related by other if ``repetitive''. Here, the algorithm searches for the potential of each metric to serve as a predictor. In some cases, one may need two well correlated metrics to aid in that decision process, while not in others. By including the results shown in the new Table 7, instead of trying to understand these correlations, we can see what metrics become relevant in the absence of others. The results indicate that in addition to the energy and the response spectra, the next in line are the peak response in acceleration, velocity, and the Arias intensity. They also confirm the lack of relevance of the duration (C11), the time integrals (C1) and (C2), and the cross correlation (C10).
}

@@

\comment{10}{%
Line 236: Remove ``A matter we discuss later.''
}

\response{%
We removed it, and in its place added an explanation about overfitting in response to the previous comment.
}

\comment{11}{%
Lines 241--243: It is not helpful to list different algorithm names for making decision trees, without any information about them or how they differ.
}

\response{%
Although it could certainly be good to have a comprehensive discussion about different methods, that would take much space, and it is not within the scope of this study. Instead, we provide adequate references to suggest further reading for the interested audience. That being said, in attention to the Reviewer's comment, we modified slightly the descritpion to give some additional context in what regards to the C5.0 algorithm and added a sentence explaining that detailing the alternative algorithms is out of the scope of this study.
}

@@

\comment{12}{%
Lines 257--258: I don't understand this sentence. What is a continuous attribute?
}

\response{%
Following Equation (3) we define $A[a]$ as all the possible values of attribute $A$. The use of $[a]$ instead of $(a)$ implies that $A$ is considered to be discrete. We therefore meant to say that if $A$ was continuous, as in $A(a)$, then the reader should refer to Quinlan (1996). We improved the text to make this clearer.
}

### 2018-04-28 abstract.tex
ricardo finished his review

% 
It has become common practice to validate ground motion simulations based on a variety of time and frequency metrics scaled to quantify the level of agreement between synthetics and data, or with respect to reference solutions. There is, however, no agreement about the importance or weight that it ought to be given to each metric. This leads to their selection often being subjective, either based on intended applications or personal preferences. As a consequence, it is difficult for simulators to identify what modeling improvements are needed, which would be easier if they could \change{focus} on a reduced number of metrics. We present an analysis \change{that looks into eleven} ground motion validation metrics using semi-supervised and supervised learning techniques\change{. These techniques} help label and classify goodness-of-fit results with the objective \change{of} prioritizing and narrowing the choice of \change{these metrics. In particular, we use the validation datasets of a series of physics-based ground motion simulations done for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake using different velocity models. We} study the relationships that exist between \change{the eleven metrics, and carry out a process where} these metrics are understood as part of a multi-dimensional space. We use a constrained \kmeans{} method, and conduct a subspace clustering analysis to address high-dimensional effects. This allows us to label a given validation dataset into four categories (poor, fair, good, excellent) following \change{previous studies}. We then develop a family of decision trees using the C5.0 algorithm, from which we select a few trees that help narrow the number of metrics leading to a validation prediction into the four referenced categories. These decision trees can be understood as rapid predictors of the quality of a simulation, \change{or as data-informed classifiers} that can help prioritize validation metrics. Our analysis\change{, although limited to the particular datasets used here,} indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of velocity and acceleration.

>>>>

%
It has become common practice to validate ground motion simulations based on a variety of time and frequency metrics scaled to quantify the level of agreement between synthetics and data or other reference solutions. There is, however, no agreement about the importance or weight that it ought to be given to each metric. This leads to their selection often being subjective, either based on intended applications or personal preferences. As a consequence, it is difficult for simulators to identify what modeling improvements are needed, which would be easier if they could \change{focus} on a reduced number of metrics. We present an analysis \change{that looks into eleven} ground motion validation metrics using semi-supervised and supervised machine learning techniques\change{. These techniques} help label and classify goodness-of-fit results with the objective \change{of} prioritizing and narrowing the choice of \change{these metrics. In particular, we use a validation dataset of a series of physics-based ground motion simulations done for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake. We} study the relationships that exist between \change{the eleven metrics, and carry out a process where} these metrics are understood as part of a multi-dimensional space. We use a constrained \kmeans{} method, and conduct a subspace clustering analysis to address the implicit high-dimensional effects. This allows us to label the data in out dataset into four validation categories (poor, fair, good, excellent) following \change{previous studies}. We then develop a family of decision trees using the C5.0 algorithm, from which we select a few trees that help narrow the number of metrics leading to a validation prediction into the four referenced categories. These decision trees can be understood as rapid predictors of the quality of a simulation, \change{or as data-informed classifiers} that can help prioritize validation metrics. Our analysis\change{, although limited to the particular dataset used here,} indicates that among the eleven metrics considered, the acceleration response spectra and total energy of velocity are the most dominant ones, followed by the peak ground response in terms of acceleration and velocity.

### 2018-04-28 dataset.tex
ricardo finished his review

We select the simulation results and validation analysis done by \citet{Taborda_2014_BSSA} as our reference dataset, hereafter the dataset. In that study, the authors carried out deterministic simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake using a finite-element approach. The simulations, done for a kinematic finite-fault model of the earthquake, were computed for a maximum frequency \fmaxeq{4} and a minimum shear wave velocity \vsmineq{200}. In total, \citet{Taborda_2014_BSSA} did three simulations, each for a different velocity model \citep[CVM-S4, CVM-H, CVM-H+GTL, see][]{Small_2017_SRL}. The modeling domain covered an area of \adomain{180}{135}{km} that included all the major sedimentary basins and other relevant geologic structures in the greater Los Angeles region. Their validation analysis consisted of comparisons with data recorded during the event at 336 ground motion monitoring stations, for the three \change{components} of motion (EW, NS, and UD). The simulation domain and the stations used for the analysis are shown in Figure \ref{fig:chino-hills}, and a sample of the validation results obtained by \citet{Taborda_2014_BSSA} using \citeauthor{Anderson_2004_Proc}'s approach is shown in Figure \ref{fig:ref-gof-maps}. This second figure, in particular, shows the spatial distribution of GOF scores (interpolated from the values at each station) for the comparison of the broadband synthetics and data, averaged for all metrics, and the three \change{components} of motion.

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/figure-02}
    \caption{Validation results obtained by \citet{Taborda_2014_BSSA} in the form of GOF values across the region of interest obtained from comparisons between simulations for three different velocity models and data recorded at ground motion monitoring stations. The color version of this figure is available only in the electronic edition.}
    \label{fig:ref-gof-maps}
\end{figure*}
% 
\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.48\textwidth]{figures/pdf/figure-03}
    \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and \change{components of motion} (NS, EW, UD). In each case, \change{the boxes represent the interquartile range ($\mathrm{IQR} = \mathrm{Q}3 - \mathrm{Q}1$), the medians are indicated by a notch in the boxes, and the vertical lines show} the range of the data, with outliers (data less than $\mathrm{Q}1 - 1.5 \mathrm{IQR}$ and greater than $\mathrm{Q}3 + 1.5 \mathrm{IQR}$) shown as scattered dots. The color version of this figure is available only in the electronic edition.}
    \label{fig:data-box-plot}
\end{figure}

Here, we use the GOF scores obtained by \citet{Taborda_2014_BSSA} independently of the velocity models and/or the \change{components of motion}. Although at times we will make distinctions between the models and the components for visualization purposes, the clustering analysis to be described in the following section was done using the whole dataset of scores. The motivation behind this choice was that the dataset, as a sample of GOF values, was independent of the simulation and serves here as a generic set for the purpose of identifying the correlations that exist between the different metrics in Table \ref{tab:metrics}. As such, given the simulations for each velocity model (3), the motion components (3), and the number of stations used in the validation (336) gave us a large enough sample of 3,024 GOF scores. Figure \ref{fig:data-box-plot} illustrates this idea by comparing the statistical distribution of the broadband GOF scores of the simulations classified by velocity models and components. It is clear that although there are differences between them, these are negligible; in other words, the statistical distribution of results for each metric is about the same independently of model or component. Consequently, we combine all the data into a single dataset.

>>>>

We select the simulation results and validation analysis done by \citet{Taborda_2014_BSSA} as our \change{\textit{validation dataset}}. In that study, the authors carried out deterministic simulations for the 2008 \eqmag{w} 5.4 Chino Hills, California, earthquake using a finite-element approach. The simulations, done for a kinematic finite-fault model of the earthquake, were computed for a maximum frequency \fmaxeq{4} and a minimum shear wave velocity \vsmineq{200}. In total, \citet{Taborda_2014_BSSA} did three simulations, each for a different velocity model \citep[CVM-S4, CVM-H, CVM-H+GTL, see][]{Small_2017_SRL}. The modeling domain covered an area of \adomain{180}{135}{km} that included all the major sedimentary basins and other relevant geologic structures in the greater Los Angeles region. Their validation analysis consisted of comparisons with data recorded during the event at 336 ground motion monitoring stations, for the three \change{components} of motion (EW, NS, and UD). The simulation domain and the stations used in that study are shown in Figure \ref{fig:chino-hills}, and a sample of the validation results obtained by the authors using \citeauthor{Anderson_2004_Proc}'s approach is shown in Figure \ref{fig:ref-gof-maps}. This second figure, in particular, shows the spatial distribution of GOF scores (interpolated from the values at each station) for the comparison of the broadband synthetics and data, averaged for all metrics and the three \change{components} of motion.

\begin{figure*}[ht!]
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/figure-01}
    \caption{\textbf{(a)} Region of interest used by \citet{Taborda_2014_BSSA} for the simulations of the 2008 $M_W ~ 5.4$ Chino Hills earthquake, including the epicenter, focal mechanism, and major quaternary faults. In the background, the main roads and topography are shown for visual reference. \textbf{(b)} Distribution of the 336 stations (gray dots) used by \citet{Taborda_2014_BSSA} for the validation analysis of their simulations within the modeling domain shown in part (a). Roads, city names, and the hillshade topography are also shown here in the background for reference. The color version of this figure is available only in the electronic edition.}
    \label{fig:chino-hills}
\end{figure*}

\begin{figure*}[t]
    \centering
    \includegraphics[width=\textwidth]{figures/pdf/figure-02}
    \caption{Validation results obtained by \citet{Taborda_2014_BSSA} in the form of GOF values across the region of interest obtained from comparisons between simulations for three different velocity models and data recorded at ground motion monitoring stations. The color version of this figure is available only in the electronic edition.}
    \label{fig:ref-gof-maps}
\end{figure*}

Here, we use the GOF scores obtained by \citet{Taborda_2014_BSSA} independently of the velocity models and/or the \change{components of motion}. \change{We refer to those scores as the \textit{data samples} in our dataset.} Although at times we will make distinctions between the models and the components for visualization purposes, the clustering analysis to be described in the following section was done using \change{all the data samples (i.e., GOF scores) in the validation dataset}. The motivation behind this choice was that the dataset, as a collection of GOF values, was independent of the simulations and serves here as a generic set of data samples for the purpose of identifying the correlations that exist between the different metrics in Table \ref{tab:metrics}. As such, given the simulations for each velocity model (3), the components of motion (3), and the number of stations used in the validation (336), gave us a large enough \change{dataset, with} 3,024 \change{data samples}. Figure \ref{fig:data-box-plot} illustrates the idea of a lack of dependence on the simulations by comparing the statistical distribution of the GOF scores of the simulations organized by velocity models and components. It is clear that although there are some differences between them, these are negligible. In other words, the statistical distribution of the \change{data samples} for each metric is about the same independently of model or component. Consequently, we felt it was alright to combine all the \change{data samples into a single dataset}.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.48\textwidth]{figures/pdf/figure-03}
    \caption{Statistical distribution of the GOF dataset obtained from \citet{Taborda_2014_BSSA} shown in the form of box-plots for each metric (C1 through C11, see Table \ref{tab:metrics}), velocity model (CVM-S, CVM-H, and CVM-H+GTL), and \change{components of motion} (NS, EW, UD). In each case, \change{the boxes represent the interquartile range ($\mathrm{IQR} = \mathrm{Q}3 - \mathrm{Q}1$), the medians are indicated by a notch in the boxes, and the vertical lines show} the range of the data, with outliers (data less than $\mathrm{Q}1 - 1.5 \mathrm{IQR}$ and greater than $\mathrm{Q}3 + 1.5 \mathrm{IQR}$) shown as scattered dots. The color version of this figure is available only in the electronic edition.}
    \label{fig:data-box-plot}
\end{figure}
